\documentclass{aa}

\usepackage{preamble}
\usepackage{newcommands}

\begin{document} 
   \title{My paper on the cosmic microwave background and formation of structures in our Universe}

   \author{V. A. Vikenes\inst{1}}
 
   \institute{Institute of Theoretical Astrophysics,  
                University of Oslo,  0315 Oslo,  Norway\\
              \email{v.a.vikenes@astro.uio.no}
             }

   \date{\today}

  \abstract{The code for this project can be found on my GitHub repository: \url{https://github.com/Vikenes/AST5220/}}

   \keywords{   cosmic microwave background  --
                large-scale structure of Universe
               }

   \maketitle

\section{Introduction} 
In this report our goal is to compute the angular power spectrum of the cosmic microwave background (CMB) radiation. In \secref{sec:M1} we compute the background cosmology, determining the evolution of the Universe in general. \secref{sec:M2} focuses on the behaviour of particles in the Universe, mainly focusing on production of neutral Hydrogen. In \secref{sec:M3} we consider cosmological perturbation theory, and study how tiny perturbation of the content of the Universe evolves on different scales. Finally, in \secref{sec:M4}, we use the perturbation quantities to compute the CMB power spectrum.  

Throughout this paper, we will use cosmological parameters from the latest Planck data {Planck2020}, unless otherwise mentioned,
\section{Fiducial parameters} \label{app:M1:fiducial_parameters}
The parameter values we use in this report are 
\begin{equation} \label{eq:Appendix:Fiducial_cosmology_parameters}
      \begin{split}
            h &= 0.67, \\ 
            \tcmb &= 2.7255\unit{K}, \\
            \neff &= 3.046, \\
            \obn &= 0.05, \\
            \ocdmn &= 0.267,\\
            \okn &=0,\\
            \onun &= \neff \cdot \frac{7}{8} \pclosed{\frac{4}{11}}^{4/3} \ogn, \\
            \oln &= 1 - (\okn + \obn + \ocdmn + \ogn + \onun), \\
            n_s &= 0.965,\\
            A_s &= 2.1\cdot 10^{-9},\\
            Y_p &= 0.245, \\
            z_\mathrm{reion} &= 8, \\
            \Delta z_\mathrm{reion} &= 0.5, \\
            z_\mathrm{Hereion} &= 3.5, \\
            \Delta z_\mathrm{Hereion} &= 0.5, \\
      \end{split}
\end{equation}   


%==============================================
%     MILESTONE 1
%==============================================
\input{../milestone1/m1tex/milestone1.tex}


%==============================================
%     MILESTONE 2
%==============================================
\input{../milestone2/m2tex/milestone2.tex}


%==============================================
%     MILESTONE 3
%==============================================
\input{../milestone3/m3tex/milestone3.tex}


%==============================================
%     MILESTONE 4
%==============================================
\input{../milestone4/m4tex/milestone4.tex}

\section{Conclusions}

Write a short summary and conclusion in the end. 

\begin{acknowledgements}
      I would like to thank Soria Moria trafikkskole for their patience regarding my lack of driving lessons, allowing me to focus on this report. I would also like to thank my mom for helping me move today (1st of June) as I eventually failed the practical driving test, following the forementioned lack of practice.     
\end{acknowledgements}


\clearpage
% \bibliographystyle{aa}
\bibliography{ref}
% \bibliographystyle{abbrvnat}
\bibliographystyle{aasjournal}
% \bibliographystyle{unsrtnat}


\end{document}
