\section{Milestone II}\label{M2}
Having successfully implemented the background cosmology, the next step in 
developing our model is to include interactions between particles. After the Big Bang, the early universe was highly ionized. Due to Thompson scattering, photons were strongly coupled to baryons. As the Universe expanded, and the temperature dropped, neutral atoms were able to form, and the photons were able to escape from the plasma. These are the CMB photons we observe today. The period where neutral atoms formed is called recombination, and will be the main topic of this section. Our goal is to compute the number density of free electrons in the Universe, and use this to estimate when recombination occurred. The evolution of free electrons will affect both structure formation and the resulting power spectrum of the CMB photons, as we will study later.

We will use both the Saha and Peebles equation to compute the electron number density, and use this to compute the optical depth. From the optical depth, we will compute the so-called \textit{visibility function}. In this section we will adapt natural units, where we set $c=\hbar=k_B=1$. Additionally, we will make the assumption that Hydrogen is the only element present in the universe. Hence, we use $Y_p=0$, rather than the value from Planck, given in 
\Eqref{eq:Appendix:Fiducial_cosmology_parameters}, as we neglect Helium and heavier elements. \cite{Dodelson}


The code for this milestone can be found on my GitHub repository: \url{https://github.com/Vikenes/AST5220/tree/main/projects/milestone2}

\input{../milestone2/m2tex/theory.tex}

\input{../milestone2/m2tex/implementations.tex}

\input{../milestone2/m2tex/results.tex} 