
\subsection{Implementation details}\label{ssec:M1:implementations} 
Something about the numerical work.

\subsubsection{Actual computations \note{(Working title)}}
The cosmological parameters we use assume $\okn=0$. In \secref{sssec:M1:implementations:supernova_fitting} we discuss how we will use supernova data to estimate a value for $\okn$. We therefore implement all functions in \secref{ssec:M1:theory} in their full form, and set $\okn=0$ for the fiducial simulations.  

\subsubsection{Splining and data analysis \note{(Working title)}}
To solve the differential equations for $\eta$ and $t$ (Eq. \eqref{eq:M1:theory:eta_ODE} and \eqref{eq:M1:theory:cosmic_time_ODE}) we use the \texttt{C++} library \texttt{GSL} \cite{gough2009gnu}, and use their Runge-Kutta4 solver. From the solution we create a spline of the results for the given $x$ domain we have considered. 

We will consider three different configurations for $x$-values. For the initial testing, we will use $x_i\in[\ln(10^{-10}),\,5]$. For the supernova data, we will use $x_i\in[\ln 10^{-2},\,0]$. When we want to estimate important times during the cosmic evolution, we will consider $x_i\in[-10,\,1]$. In all cases, we use $N_x=10^5$ number of points.  

\note{Make new simulations...} 


\subsubsection{MCMC \note{(Working title)}}\label{sssec:M1:implementations:supernova_fitting}
The supernova data we will use contains $N=31$ data points of luminosity distance, $d_L^\mathrm{obs}(z_i)$, with associated measurement errors, $\sigma_i$, at different redshifts, $z_i\in[0.01,\,1.30]$. This approximately corresponds to $x_i\in[-9.95\cdot 10^{-3},\,-0.833]$. Using these measurements, we want to constrain the three-dimensional parameter space 
\begin{equation}
    \mathcal{C} = \Big\{ \hest,\: \omest,\: \okest \Big\}, \label{eq:M1:implementations:supernova_parameter_space}
\end{equation}  
where the hat is used to distinguish them from the fiducial values. \note{Fiducial?}. We use $\obn=0.05$ for this analysis, so $\omest$ enters via $\ocdmn=\omest-\obn$. Additionally, the neutrinos are not relevant in the regime we consider here, so we also set $N_\mathrm{eff}=0$.  

We will assume that the measurements at different redshifts are normal distributed and uncorrelated. The likelihood function is then given by $L\propto e^{-\chi^2/2}$, where 
\begin{equation}
    \chi^2(\C) = \sum_{i=1}^N \frac{[d_L(z_i, \C) - d_L^\mathrm{obs}(z_i)]^2}{\sigma_i^2}, \label{eq:M1:implementations:chi2_supernova_data}
\end{equation}
is the function we want to minimize. \note{Little h needs to be defined.} To do this, we will sample parameter values randomly by a MCMC process. The parameter space we will explore is limited to  
\begin{equation} \label{eq:M1:implementations:supernova_parameter_ranges}
    \begin{split}
        0.5 < \hest < 1.5, \\
        0 < \omest < 1, \\
        -1 < \okest < 1.
    \end{split}
\end{equation}   
To generate a new sample, we update each parameter by generating a random number $P\sim\mathcal{N}(0,1)$, and multiplying it by a step size. We will use step sizes of $\Delta \hest = 0.007,\,\Delta \omest=0.05,\,\Delta\okest=0.05$. To determine whether a new configuration should be included in the sample we use the Metropolis algorithm, where we always accept a state if it yields a lower value of $\chi^2$ compared to the previous state that was accepted. If the new value of $\chi^2$ is greater than the old one, we accept it if the ratio of the likelihood functions $L(\chi^2_\mathrm{new})/L(\chi^2_\mathrm{old})>p$, where $p\sim\mathcal{U}(0,1)$. We continue drawing samples until we get a total of $\hat{n}=10^4$ samples. For the samples generated, we omit the first $1000$ samples of the chain from our analysis.   

With our generated samples, we can use the best fit, $\chi^2_\mathrm{min}$, to find the $1\sigma$ and $2\sigma$ confidence regions. For the $\chi^2$ distribution with $3$ parameters, these regions are given by $\chi^2 - \chi^2_\mathrm{min}<3.53$ and $\chi^2 - \chi^2_\mathrm{min}<8.02$, respectively. We will plot the $1\sigma$ and $2\sigma$ constraint in the $(\omn,\oln)$ plane. Since $\oradn<10^{-4}$, $\oln$ can be approximated well by $\oln=1-\omn$. After that we will plot the posterior probability distribution function (PDF) for $H_0$. 

To compare our fit with the Planck data, we will plot $d_L^\mathrm{obs}(z_i)$ together with $d_L^\mathrm{fit}(z)$ and $d_L^\mathrm{Planck}(z)$. We do this by computing the background with $h,\,\omn,\,\okn$ replaced with the values of $\hest,\,\omest,\,\okest$ corresponding to $\chi^2_\mathrm{min}$.     



