\section{Milestone I}\label{M1}
In this section we will examine the evolution of the Universe's uniform background. Our primary objective is to develop methods for computing the Hubble parameter and related time- and distance measures. These methods provide a first step towards further investigations and modelling of the early Universe. To compute the background cosmology, we will solve ordinary differential equations (ODEs) numerically, using cosmological parameters obtained from the Planck Collaboration \cite*{Planck2020}. The parameters we will use are listed in Eq. \eqref{eq:Appendix:Fiducial_cosmology_parameters} in  \appref{app:M1:fiducial_parameters}. One crucial aspect in the process is validating our model. We will therefore develop some simple methods for comparing our result. This will mainly involve considering simplified cases where analytical solutions can be obtained. 

Our primary focus in this section concerns methods where the cosmological parameters are given from the start. Another interesting aspect is to use data to constrain cosmological parameters. To do this, we will use data from supernova observations \cite{Supernova2014Betoule}, containing luminosity distance associated with different values of redshift. By employing the numerical methods we develop initially, we will try to estimate optimal values of three cosmological parameters, by implement a simple Markov chain Monte Carlo (MCMC) algorithm. The parameters we will be sampling are $h,\,\omn$ and $\okn$. From these results, we will investigate confidence regions of $\omn$ and $\oln$, and try to estimate a probability distribution function (PDF) for the Hubble parameter. 


\input{../milestone1/report/theory.tex}

\input{../milestone1/report/implementations.tex}

\input{../milestone1/report/results.tex}